---
title: "svyEffects"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

```{r, echo = FALSE}
library(tidyverse)
library(ggplot2)
```


## Introduction

An oft-cited reason why `R` is not more widely used in social science research is its disjointed and incomplete set of tools to deal with weights. `{svyEffects}` helps address this problem by providing a suite of post-estimation tools for working with limited dependent variable models (binary, ordinal, and multinomial logit) estimated on survey-weighted data.

Its main set of functions calculate predicted probabilities using either:

- the *average marginal effects* approach (also known as *marginal effects at observed values*, or *adjusted predictions*); or 
- the *marginal effects at reasonable/representative/typical values* approach (also known as *marginal effects for the average case*). 

These approaches are analogous to Stata's commands `margins x` and `margins x, at`, respectively.

After calculating predicted probabilities, it will then calculate differences in probabilities (also known as *contrasts*/*pairwise comparisons* for categorical predictors or *first differences* for continuous predictors) using:

- for continuous predictors, the change across the entire range of the variable (by default), or a one-unit or one-standard-deviation change centred on the mean; or
- for categorical predictors, all pairwise differences.

For both predictions and differences, it uses simulation methods (the parametric bootstrap) to derive 95% confidence intervals.

It works with the following survey-weighted models or (non-survey) weighted models (i.e. models estimated with the `weight=` option enabled):

- Binary logit models
  - `survey::svyglm`
  - `glm`
- Ordinal logit models
  - `MASS::polr`
  - `survey::svyolr`
- Multinomial logit models
  - `svrepmisc::svymultinom`
  - `nnet::multinom`

Also included in the package are:
 
- A snippet of the 2019 Canadian Election Study online survey for testing and demonstration purposes. This can be loaded with the command `data(ces19w)`.
- A `plot()` method that creates a `ggplot` object of predicted probabilities or differences in predicted probabilities. This plot can be modified by adding further `ggplot` commands, which is shown below.
- A function `mnlSig` that displays a concise summary of multinomial logit coefficients with statistical significance stars. This has been adapted for use on `svymultinom` objects from Dave Armstrong's original function from `{DAMisc}`, which works for `multinom` objects.


--------------------------------------------------------------------------------


# Development history and differences from other packages

This package extends functions originally written by Dave Armstrong, some of which are in his `{DAMisc}` package (https://github.com/davidaarmstrong/damisc). 

The reporting functions and naming conventions are inspired by Daniel Ludecke's excellent `{ggeffects}` package (https://github.com/strengejacke/ggeffects). Current users of `ggeffects` will notice similarities between `{svyEffects}` and `{ggeffects}`. However, while `{ggeffects}` can estimate MER probabilities (what it calls *adjusted predictions*) with `svyglm` objects, it is not compatible with either `svyolr` or `svymultinom` objects. Moreover, `{svyEffects}` estimates true average marginal effects, which is the estimate of a variable's effect on a given outcome at the population level as opposed to a variable's effect for a hypothetical "average case" that may or may not exist or even be theoretically plausible. (A detailed discussion of the difference is in Hanmer and Kalkan 2013, *AJPS*, the full citation of which can be found in the reference section at the end of this readme.)

Note: because AMEs run simulations on multiple copies of your dataset, they can take much more time to calculate than MERs, particularly on large datasets or when using an older computer. Those needing quick results can calculate MERs (which, in practice *usually* substantively similar to AMEs) and then decide from there for which variables they want to calculate AMEs. 


--------------------------------------------------------------------------------


# Binary dependent variable models

To demonstrate how this function works with binary dependent variables, we'll model voting for the Conservative Party of Canada versus voting for any other party.

```{r}
data(ces19w)

library(survey)
ces19w_svy <- survey::svydesign(ids = ~1, strata = NULL, weights = ~sampleweight, 
                                data = ces19w, digits = 3)

VOTECON <- survey::svyglm(votecon ~ agegrp + educ + region + market, 
                          design = ces19w_svy, family = binomial)
summary(VOTECON)
```

Let's look at the effect of educational attainment (`educ`), a categorical predictor with three levels: high school or less, some post-secondary, and a university degree at the bachelor's level or higher.

The function `svyAME` will return average marginal effects for education, or the effect of a change in education, holding all other variables at observed values. We'll specify a seed value for reproducibility purposes.

The function's output is a list that contains three data frames: 

- `$preds`: predicted probabilities
- `$diffs`: differences in predicted probabilities
- `$seed`: the seed value used for the simulations

```{r}
library(svyEffects)
VOTECON_educ_ame <- svyEffects::svyAME(VOTECON,
                                       varname = "educ",
                                       weightvar = "sampleweight",
                                       seed = 2019)
VOTECON_educ_ame$preds
VOTECON_educ_ame$diffs
VOTECON_educ_ame$seed
```

The outputs of this function lend themselves well to plotting using `{ggplot2}`. As an example, let's plot the predicted probabilities of voting Conservative across levels of education.

```{r}
library(ggplot2)
ggplot(VOTECON_educ_ame$preds) +
  aes(x = educ,
      y = predicted,
      ymin = conf.low,
      ymax = conf.high) +
  geom_pointrange() +
  labs(title = "Probability of voting Conservative by education",
       y = "Predicted probability",
       x = "Education")
```

For convenience, `{svyEffects}` also includes a `plot()` method, which uses the `{ggplot2}` engine to visualize either predicted probabilities or differences in predicted probabilities.

By default, the predicted probabilities are plotted, as shown below. 

```{r}
plot(VOTECON_educ_ame)
```

Note that labelling is minimal on the automatically-generated plots, but you can add your own customization using `{ggplot2}`'s code conventions.

```{r}
plot(VOTECON_educ_ame) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "My title",
       subtitle = "My subtitle",
       x = "My xvar label",
       y = "My yvar label",
       caption = "My caption") +
  theme_classic()
```

You can also plot the differences in predicted probabilities between levels of education by including the option `what = "diffs"` (or simply `"diffs"`) in the `plot()` function call.

```{r}
plot(VOTECON_educ_ame, "diffs")
```

Now, let's look at the effect of market liberalism, a continuous predictor that ranges from -1 (minimal market liberalism, or the most left-wing position) to +1 (maximal market liberalism, or the most right-wing position).

Note that, because the function returns a first difference for continuous predictors, the graph is not any more illuminating than the summary statistic.

```{r}
VOTECON_market_ame <- svyAME(VOTECON,
                             varname = "market",
                             weightvar = "sampleweight",
                             seed = 2019)
VOTECON_market_ame$preds
VOTECON_market_ame$diffs
VOTECON_market_ame$seed
plot(VOTECON_market_ame)
plot(VOTECON_market_ame, "diffs")
```


--------------------------------------------------------------------------------


# Ordinal dependent variable models

To demonstrate ordinal dependent variables, we'll model feeling thermometer ratings for the leader of the Conservative Party of Canada. This variable usually ranges from 0 to 100. But, for this example, we'll used a binned ordinal measure of "cold" (0-39), "lukewarm" (40-59), and "hot" (60-100). 


```{r}
data(ces19w)
ces19w <- ces19w

library(survey)
ces19w_svy <- svydesign(ids = ~1, strata = NULL, weights = ~sampleweight, 
                        data = ces19w, digits = 3)

CONLDR <- svyolr(ldr_con_ft ~ agegrp + educ + region + market, 
                 design = ces19w_svy)
summary(CONLDR)
```

Here's the effect of education on feelings towards the Conservative Party leader.

```{r}
CONLDR_educ_ame <- svyAME(CONLDR,
                          varname = "educ",
                          weightvar = "sampleweight",
                          seed = 2019)
CONLDR_educ_ame$preds
CONLDR_educ_ame$diffs
plot(CONLDR_educ_ame)
plot(CONLDR_educ_ame, "diffs")
```

Here's the effect of market liberalism.

```{r}
CONLDR_market_ame <- svyAME(CONLDR,
                            varname = "market",
                            weightvar = "sampleweight",
                            diffchange = "range",
                            seed = 2019)
CONLDR_market_ame$preds
CONLDR_market_ame$diffs
plot(CONLDR_market_ame)
plot(CONLDR_market_ame, "diffs")
```

For ordinal and multinomial probabilities, the plot method follows the conventions used by the `{ggeffects}` package (i.e. facetting by response level). But, you can re-create the `Stata` default of colour-coding the response level, by writing your own `ggplot` command, as shown below.

```{r}
ggplot(CONLDR_market_ame$preds) +
  aes(x = market, y = predicted, ymin = conf.low, ymax = conf.high, colour = y, fill = y) +
  geom_line() +
  geom_ribbon(colour = "transparent", alpha = 0.2) +
  labs(title = "Effect of market liberalism on Conservative leader ratings",
       x = "Market liberalism (least to most)",
       y = "Predicted probability",
       fill = "Rating",
       colour = "Rating") +
  theme_bw()
```

--------------------------------------------------------------------------------


# Multinomial dependent variable models

To demonstrate multinomial dependent variables, we'll model vote choice in the 2019 Canadian Federal Election. To keep things simple, we'll limit our analysis to the three major parties (the Liberals, Conservatives, and New Democrats) and exclude the province of Quebec (which has a different party system and patterns of vote choice).

There is no way to directly estimate a multinomial model with the `{survey}` package in R. The package `{svyrepmisc}` generates an approximation by turning the weighting scheme into replicate weights and estimating the model with those. It uses the jackknife to calculate variances.

We'll go through this process step-by-step. First, we'll import the data, do some data cleaning, and then create our usual survey-design object.

```{r}
data(ces19w)
ces19w <- ces19w %>%
  filter(region != "Quebec") %>%
  filter(vote %in% c("Liberal", "Conservative", "NDP")) %>%
  mutate(region = droplevels(region),
         vote = droplevels(vote)) 

library(survey)
ces19w_svy <- svydesign(ids = ~1, strata = NULL, weights = ~sampleweight, 
                        data = ces19w, digits = 3)
```

Now, we'll use the function `as.svrepdesign()` from `{survey}` to turn our sampling weights into replicate weights with variances calculated using the jackknife.

```{r}
ces19w_svy_r <- as.svrepdesign(ces19w_svy, type = "JK1")
```


After our survey design object with replicate weights and jackknife variances is created, we can use the function `svymultinom` from `{svyrepmisc}` to run our vote choice model. 

Note: use the option `trace = FALSE` in the `svymultinom()` function call to suppress the reporting of each replication (similar to using the option `quietly` in Stata).

Included with `{svyEffects}` the function `mnlSig`, which displays coefficients from multinomial logit models and flags statistically significant ones. `mnlSig` is adapted from Dave Armstrong's original function from his `{DAMisc}` package.


```{r}
# remotes::install_github("carlganz/svrepmisc")
library(svrepmisc)

VOTE <- svymultinom(vote ~ region + educ + relig + market, 
                    design = ces19w_svy_r, trace = FALSE)
mnlSig(VOTE)
```

For our post-estimation command, we'll need to specify a few more options because `svymultinom` does not store them in it's output. These are:

- `design = `: the survey design object used to estimate the model; and
- `modform = `: the model formula used in the `svymultinom` call.

Here's the effect of education.

```{r}
VOTE_educ_ame <- svyAME(VOTE,
                        varname = "educ",
                        weightvar = "sampleweight",
                        seed = 2019,
                        design = ces19w_svy_r,
                        modform = "vote ~ region + educ + relig + market")
VOTE_educ_ame$preds
VOTE_educ_ame$diffs
plot(VOTE_educ_ame)
plot(VOTE_educ_ame, "diffs")
```

Here's the effect of market liberalism.

```{r}
VOTE_market_ame <- svyAME(VOTE,
                          varname = "market",
                          weightvar = "sampleweight",
                          design = ces19w_svy_r,
                          diffchange = "range",
                          modform = "vote ~ region + educ + relig + market",
                          seed = 2019)
VOTE_market_ame$preds
VOTE_market_ame$diffs
plot(VOTE_market_ame)
plot(VOTE_market_ame, "diffs")
```


--------------------------------------------------------------------------------


# Planned features

This package is under active development, and several features will be added, 
including:

- support for interaction terms (interaction terms in MER probabilities is coming
in the next major update; interaction in AME probabilities is still under development).
- support for using an alternative variance-covariance matrix with the `sandwich` 
package (note: this is only for binary logit models because `sandwich` does not
play nice with ordinal or multinomial models; that said, survey-weighted models 
do adjust the variance-covariance matrix, so this is only useful if you really 
like one adjustment method in particular).
- more documentation and in-depth comparison to results from Stata.


--------------------------------------------------------------------------------


# References

Hanmer, M.J. and K.O. Kalkan.  2013. "Behind the Curve: Clarifying the Best 
Approach to Calculating Predicted Probabilities and Marginal Effects from Limited 
Dependent Variable Models." _American Journal of Political Science_. 57(1): 263-277.

Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, 
"2019 Canadian Election Study - Online Survey," https://doi.org/10.7910/DVN/DUS88V, 
Harvard Dataverse, V1.

